Efficient Calculation of Improved Simultaneous Testing Bands for QQplots with \proglang{R} Package \pkg{qqconf}

### 2021-01-24 article/manuscript.Rnw
adding article

\author{Eric Weine\\University of Chicago
   \And Mary Sara McPeek\\University of Chicago}
\Plainauthor{Eric Weine, Mary Sara McPeek}

\title{Efficient Calculation of Improved Simultaneous Testing Bands for QQplots}
%\author{Eric Weine and Mary Sara McPeek}
\Abstract{
Quantile-Quantile plots (QQplots) are often difficult to interpret because it is unclear how large the deviation from the theoretical distribution must be to indicate a lack of fit. Current packages and algorithms towards this end either do not ensure a robust Type I error rate, are too slow, or are under-powered to deviations in the tails of the distribution. In this paper, we present an efficient algorithm that computes simultaneous testing bands for QQplots. These bands have a variety of desirable properties, including being fast to compute and being equally sensitive to deviations in all parts of the null distribution, including the tails.
}

@@

\Address{
  Mary Sara McPeek\\
  Department of Statistics\\
  University of Chicago\\
  924 E 57th St., Chicago, IL\\
  E-mail: \email{mcpeek@uchicago.edu}\\
  URL: \url{https://galton.uchicago.edu/~mcpeek/}
}
\begin{document}

@@

\section{Introduction}
Quantile-Quantile plots (QQplots) are a common statistical tool used to judge if a sample comes from a specified distribution \citep{wilk_gnanadesikan_1968}. Despite their ubiquity, they are often difficult to interpret because it is challenging to determine if the magnitude of the deviation from the specified distribution is large enough to indicate a lack of fit as opposed to sampling variance. To make this determination, it is useful to put testing bands on a QQplot.
\newline
\newline
A number of methods have been created towards this end. First, some available software puts pointwise testing bands onto QQplots, which conduct $\alpha$ level tests on each order statistic \citep{qqplotr}. This method is clearly insufficient, as it ignores the multiple testing problem and thus provides no guarantee on the Type I error rate of testing the global null hypothesis that all of the data come from the specified distribution. Second, the Komolgorov-Smirnov (KS) test is a common method for creating joint bands \citep{kolmogoroff1941confidence, smirnov1944approximate}. While this method ensures the correct Type I error by using the distribution of the KS statistic, this test suffers from very low power under a variety of reasonable alternatives \citep{aldor2013power}. In the same work, Aldor-Noiman et al. introduced a much more powerful test (under most alternatives) called the Tail Sensitive (TS) test. This test uses simulation to conduct an $\eta$ level test on each order statistic such that the overall test has an $\alpha$ level Type I error. This test has a variety of desirable properties, but because it requires simulation it is extremely slow in large datasets.
\newline
\newline
In what follows, we present a method that yields the same results as the TS test but does so substantially faster because of an efficient recursive calculation. We then demonstrate this test on a few multiple testing scenarios including Type I error calibration and genetic studies.
\section{Methods}

@@

\begin{equation*}
  X_{1}, ..., X_{n} \distras{iid} F,
\end{equation*}
and we are interested in conducting the following hypothesis test

\begin{equation*}
      H_{0}: F = F_{0} \textrm{ vs. } H_{A}: F \neq F_{0}
\end{equation*}
with level $\alpha$, where all parameters of $F_{0}$ are known. One approach to this problem, referred to as ``local levels,'' is to conduct separate hypothesis tests on each of the order statistics $X_{(i)}$, where each test has Type I error rate $\eta_{i}$ \citep{gontscharuk2016goodness}. Then, we reject the global null hypothesis if at least one of the $n$ tests results in a rejection. That is, we construct a set of intervals

\begin{equation*}
    (h_{1}, g_{1}), ..., (h_{n}, g_{n}),
\end{equation*}
and we reject $H_{0}$ if for any $i$

\begin{equation*}
    X_{(i)} \not\in (h_{i}, g_{i}).
\end{equation*}
If we know the local level $\eta_{i}$, then
\begin{equation}
    h_{i} = F^{-1}_{i}(\eta_{i} / 2) \text{ and } g_{i} = F^{-1}_{i}(1 - \eta_{i} / 2),
\end{equation}
where $F_{i}$ is the CDF of the $i^{th}$ order statistic. Thus, the difficulty is in determining an appropriate vector of local levels $(\eta_{1}, ..., \eta_{n})$ for which the resulting global level is $\alpha$. In our case, we want to create testing bands that are ``agnostic'' to any alternative distribution. By this, we mean that we would like to design a test that applies equal scrutiny to each order statistic, and thus we set
% an alternative explanation is that we're trying to design a test that would have optimum power under the alternative distribution where the probability of any of the order statistics being outside of the pointwise null region is equal.
\begin{equation}
    \eta_{1} = \eta_{2} = ... = \eta_{n} = \eta.
\end{equation}

This is known as ``equal local levels.''
\newline
\newline
\noindent Given the sample size $n$ and the desired level $\alpha$ we present the following algorithm for obtaining $\eta$ in two steps:
\newline
\newline
1. Given a proposed testing region $(h_{1}, g_{1}), ..., (h_{n}, g_{n})$, where $h_{i} < g_{i}$ for $i = 1, ..., n$, find the probability of $X_{(1)}, ..., X_{(n)}$ falling outside of this region under the null.
\newline
\newline
2. Using the algorithm described in step 1, conduct a binary search on the space of the local level $\eta$. This involves proposing a value of $\eta$, calculating the corresponding region $(h_{1}, g_{1}), ..., (h_{n}, g_{n})$ using equation (1), and then using step 1 to check if the region achieves the desired global Type I error rate $\alpha$.
\newline
\newline
We now describe these two steps in more detail:
\newline
\newline
\textbf{Step 1:} First, note that under the null hypothesis,

\begin{equation*}
    F_{0}(X_{1}), ..., F_{0}(X_{n}) \distras{iid} U(0, 1).
\end{equation*}
Thus, without loss of generality, we can build the algorithm for step 1 assuming that

\begin{equation*}
    X_{1}, ..., X_{n} \distras{iid} U(0, 1)
\end{equation*}
% I'll worry about the uncertainty in parameter estimates later
% I'll probably want some sort of a diagram here eventually
We are trying to calculate the following probability
\begin{equation*}
    \alpha = P_{0}\Big(\bigcup\limits_{i=1}^{n} X_{(i)} \notin (h_{i}, g_{i})\Big)
\end{equation*}
Where $P_{0}$ is the probability under the null hypothesis. It is easier to calculate this as
\begin{equation*}
    \alpha = 1 - P_{0}\Big(\bigcap\limits_{i=1}^{n} X_{(i)} \in (h_{i}, g_{i})\Big)
\end{equation*}
\newline
To calculate this value, it is easier to transform the events over order statistics to multinomial events. To do this, we introduce the following notation:
\newline
\newline
Let $b_{1}, ..., b_{2n}$ be the sorted values of $h_{1}, ..., h_{n}, g_{1}, ..., g_{n}$ in ascending order. We also define $b_{0} = 0$ and $b_{2n + 1} = 1$. We will divide the interval $(b_{0}, b_{2n + 1})$ into $2n + 1$ bins, where bin 1 is $(b_{0}, b_{1})$, bin 2 is $(b_{1}, b_{2}), ...,$ and bin $2n + 1$ is $(b_{2n}, b_{2n + 1})$. Note that because the uniform distribution is continuous, the probabilities associated with the endpoints of the bins are 0, so it does not matter if the left and right boundaries of these intervals are open or closed. Let $N_{j} = \sum_{i = 1}^{n}\mathds{1}\Big(X_{i} \in (b_{j - 1}, b_{j})\Big)$ denote the random variable that counts the number of X's falling into bin j, for $1 \leq j \leq 2n+1$, and let $S_{k} =  \sum_{j=1}^{k} N_{j}$ be the $kth$ partial sum of the $N$'s, for $1 \leq k \leq 2n+1$. The key to the algorithm is that the following two events are the same:
\begin{equation*}
    \{X_{(i)} \in (h_{i}, g_{i}) \text{ for } i = 1, ..., n\} = \{l_{k} \leq S_{k} \leq u_{k} \textrm{ for } k = 1, 2, ..., 2n\},
\end{equation*}
where

@@

@@

@@

Empirically, it seems that our algorithm has computational complexity $O(n^{2})$. For a dense grid of values of $n$ between $10$ and $50,000$, the number of recursive steps required (line 15 in algorithm 1 above) scales with $n^{2}$. More precisely, it appears that the number of recursive steps required is very close to $8n^{2}$. While each recursive step itself requires calculating a binomial probability which is $O((n + 1)!)$ due to the calculation of the binomial coefficient multiplied by a quantity with powers as large as $n$, these calculations can be memoized with $O(n^{2})$ cost. Thus, these binomial probability calculations do not change the big $O$ complexity of the algorithm.

@@

As noted above, the algorithm presented is only correct if all of the parameters of $F_{0}$ are known, and all $X_{1}, ..., X_{n}$ are independent. While both of these assumptions may only hold in a small number of cases, the methods above are still useful in a variety of scenarios.

@@

One of the main advantages of the local levels method, is that it can easily be used to put testing bands onto QQplots by simply graphing each $(h_{i}, g_{i})$ interval. This allows us to examine how a dataset might deviate from some null distribution much better than simply applying a test that yields a binary conclusion. Below, we present a few examples where a QQplot is useful, and where the local levels test seems ideal for assessing deviation from a global null hypothesis.

@@

% It may also make sense for me to make a comparison to the Kolmogorov-Smirnov statistic at some point. I should probably include some instructive graphs or something.
% I want to have a different paragraph to better introduce the concept
% This shouldn't be too difficult
Suppose we have devised a new testing procedure to test a null hypothesis $H_{0}$ with test statistic $T(X_{1}, ..., X_{n})$, and we wish to establish that this test has correct Type I error rate.
\newline
\newline
Typically, the verification of Type I error rate is done using the following procedure:
\newline
\newline
(1) Generate $m$ simulated datasets: $(X_{1}^{(1)}, ..., X_{n}^{(1)}), ..., (X_{1}^{(m)}, ..., X_{n}^{(m)})$ under $H_{0}$.
\newline
\newline
(2) Apply the test $T$ to each of these datasets at level $\alpha$ to yield $m$ reject / not-reject conclusions: $c_{1}, ..., c_{m}$, where $c_{i} = 1$ if $H_{0}$ is rejected and $0$ otherwise.
\newline
\newline
(3) Conduct a test to confirm that the mean of $c_{1}, ..., c_{m}$ is $\alpha$.
\newline
\newline
While the above procedure provides reliable information about the Type I error calibration for one level of $\alpha$, it provides little information about the global calibration of p-values. Instead, we suggest the following procedure:
\newline
\newline
(1) As above.
\newline
\newline
(2) Apply the test $T$ to each of these datasets to yield $m$ p-values: $p_{1}, ..., p_{m}$.
\newline
\newline
(3) Apply the local levels procedure to test if $p_{1}, ..., p_{m} \distras{iid} U(0, 1)$.
% It seems to me that this would be an opportune time to make a comparison to the KS-test, but I'm not totally sure.
% I should maybe put some graphs here as well.
\newline
\newline
This allows us to easily visualize the global calibration of the p-values with just one graph and diagnose any issues if they exist. In Step 3, one could use many different testing bands. However, in the calibration of p-values, we typically don't have the expectation that our p-values would be more likely to deviate from uniform in some regions than others, and so it makes sense to use the local levels test because it is agnostic to the alternative distribution. Moreover, since it is generally most concerning if p-values are not calibrated in the lower tail of the distribution, the local levels test is preferable to the standard KS test because it is much more sensitive in the tails (Aldor-Noiman 2013).

\subsubsection{Chi-Square Test for Independence in a \texorpdfstring{$2 \times 2$}{2 X 2} Table}

\bibliography{refs}
\end{document}

>>>>

\author{Eric Weine\\University of Chicago
   \And Mary Sara McPeek\\University of Chicago
   \And Mark Abney\\University of Chicago}
\Plainauthor{Eric Weine, Mary Sara McPeek, Mark Abney}

\title{Efficient Calculation of Improved Simultaneous Testing Bands for QQplots with \proglang{R} Package \pkg{qqconf}}
\Plaintitle{Efficient Calculation of Improved Simultaneous Testing Bands for QQplots with R Package qqconf}
\Shorttitle{Efficient Calculation of Simultaneous Testing Bands for QQplots with \pkg{qqconf}}
%\author{Eric Weine and Mary Sara McPeek}
\Abstract{
  Quantile-Quantile plots (QQplots) are often difficult to interpret because it is unclear how large the deviation from the theoretical distribution must be to indicate a lack of fit. Current packages and algorithms towards this end either do not ensure a robust Type I error rate, are too slow, or are under-powered to deviations in the tails of the distribution. In this paper, we present an efficient algorithm and accompanying \proglang{R} Package \pkg{qqconf} that computes simultaneous testing bands for QQplots. These bands have a variety of desirable properties, including being fast to compute and being equally sensitive to deviations in all parts of the null distribution, including the tails.
}

@@

\Address{
  Eric Weine\\
  Department of Statistics\\
  University of Chicago\\
  924 E 57th St., Chicago, IL\\
  E-mail: \email{ericweine15@gmail.com}\\
  \\
  Mary Sara McPeek\\
  Department of Statistics\\
  University of Chicago\\
  924 E 57th St., Chicago, IL\\
  E-mail: \email{mcpeek@uchicago.edu}\\
  URL: \url{https://www.stat.uchicago.edu/~mcpeek/}\\
  \\
  Mark Abney\\
  Department of Human Genetics\\
  University of Chicago\\
  920 E 58th St., Chicago, IL\\
  E-mail: \email{abney@uchicago.edu}\\
  URL: \url{https://hgen.uchicago.edu/program/faculty/mark-abney}
}

\begin{document}

@@

\section{Introduction}
Quantile-Quantile plots (QQplots) are a common statistical tool used to judge if a sample comes from a specified distribution \citep{wilk_gnanadesikan_1968}. Despite their ubiquity, they are often difficult to interpret because it is challenging to determine if the magnitude of the deviation from the specified distribution is large enough to indicate a lack of fit as opposed to sampling variance. To make this determination, it is useful to put testing bands on a QQplot.
\newline
\newline
A number of methods have been created towards this end. First, some available software, like the \proglang{R} package \pkg{qqplotr}, puts pointwise testing bands onto QQplots, which conduct $\alpha$ level tests on each order statistic \citep{qqplotr}. This method is clearly insufficient, as it ignores the multiple testing problem and thus provides no guarantee on the Type I error rate of testing the global null hypothesis that all of the data come from the specified distribution. Second, the Komolgorov-Smirnov (KS) test, a part of the \pkg{stats} package in \proglang{R}, is a common method for creating joint bands \citep{kolmogoroff1941confidence, smirnov1944approximate}. While this method ensures the correct Type I error by using the distribution of the KS statistic, this test suffers from very low power under a variety of reasonable alternatives \citep{aldor2013power}. In the same work, Aldor-Noiman et al. introduced a much more powerful test (under most alternatives) called the Tail Sensitive (TS) test. This test uses simulation to conduct an $\eta$ level test on each order statistic such that the overall test has an $\alpha$ level Type I error. This test has a variety of desirable properties, but because it requires simulation it is extremely slow in large datasets. It is also available in the package \pkg{qqplotr}.
\newline
\newline
In what follows, we present a method that yields the same results as the TS test but does so substantially faster because of an efficient recursive calculation. We then demonstrate this test on a few multiple testing scenarios including Type I error calibration and genetic studies.
\section{Methods}

@@

%
  X_{1}, ..., X_{n} \distras{iid} F,
%
and we are interested in conducting the following hypothesis test
%
      H_{0}: F = F_{0} \textrm{ vs. } H_{A}: F \neq F_{0}
%
with level $\alpha$, where all parameters of $F_{0}$ are known. One approach to this problem, referred to as ``local levels,'' is to conduct separate hypothesis tests on each of the order statistics $X_{(i)}$, where each test has Type I error rate $\eta_{i}$ \citep{gontscharuk2016goodness}. Then, we reject the global null hypothesis if at least one of the $n$ tests results in a rejection. That is, we construct a set of intervals
%
    (h_{1}, g_{1}), ..., (h_{n}, g_{n}),
%
and we reject $H_{0}$ if for any $i$
%
    X_{(i)} \not\in (h_{i}, g_{i}).
%
If we know the local level $\eta_{i}$, then
%
    h_{i} = F^{-1}_{i}(\eta_{i} / 2) \text{ and } g_{i} = F^{-1}_{i}(1 - \eta_{i} / 2),
%
where $F_{i}$ is the CDF of the $i^{th}$ order statistic. Thus, the difficulty is in determining an appropriate vector of local levels $(\eta_{1}, ..., \eta_{n})$ for which the resulting global level is $\alpha$. In our case, we want to create testing bands that are ``agnostic'' to any alternative distribution. By this, we mean that we would like to design a test that applies equal scrutiny to each order statistic, and thus we set
% 
    \eta_{1} = \eta_{2} = ... = \eta_{n} = \eta.
%
This is known as ``equal local levels.''
\noindent Given the sample size $n$ and the desired level $\alpha$ we present the following algorithm for obtaining $\eta$ in two steps:
1. Given a proposed testing region $(h_{1}, g_{1}), ..., (h_{n}, g_{n})$, where $h_{i} < g_{i}$ for $i = 1, ..., n$, find the probability of $X_{(1)}, ..., X_{(n)}$ falling outside of this region under the null.
2. Using the algorithm described in step 1, conduct a binary search on the space of the local level $\eta$. This involves proposing a value of $\eta$, calculating the corresponding region $(h_{1}, g_{1}), ..., (h_{n}, g_{n})$ using equation (1), and then using step 1 to check if the region achieves the desired global Type I error rate $\alpha$.
We now describe these two steps in more detail:
\textbf{Step 1:} First, note that under the null hypothesis,
%
    F_{0}(X_{1}), ..., F_{0}(X_{n}) \distras{iid} U(0, 1).
%
Thus, without loss of generality, we can build the algorithm for step 1 assuming that
%
    X_{1}, ..., X_{n} \distras{iid} U(0, 1)
% 
We are trying to calculate the following probability
%
    \alpha = P_{0}\Big(\bigcup\limits_{i=1}^{n} X_{(i)} \notin (h_{i}, g_{i})\Big)
%
Where $P_{0}$ is the probability under the null hypothesis. It is easier to calculate this as
%
    \alpha = 1 - P_{0}\Big(\bigcap\limits_{i=1}^{n} X_{(i)} \in (h_{i}, g_{i})\Big)
%
To calculate this value, it is easier to transform the events over order statistics to multinomial events. To do this, we introduce the following notation:
Let $b_{1}, ..., b_{2n}$ be the sorted values of $h_{1}, ..., h_{n}, g_{1}, ..., g_{n}$ in ascending order. We also define $b_{0} = 0$ and $b_{2n + 1} = 1$. We will divide the interval $(b_{0}, b_{2n + 1})$ into $2n + 1$ bins, where bin 1 is $(b_{0}, b_{1})$, bin 2 is $(b_{1}, b_{2}), ...,$ and bin $2n + 1$ is $(b_{2n}, b_{2n + 1})$. Note that because the uniform distribution is continuous, the probabilities associated with the endpoints of the bins are 0, so it does not matter if the left and right boundaries of these intervals are open or closed. Let $N_{j} = \sum_{i = 1}^{n}\mathds{1}\Big(X_{i} \in (b_{j - 1}, b_{j})\Big)$ denote the random variable that counts the number of X's falling into bin j, for $1 \leq j \leq 2n+1$, and let $S_{k} =  \sum_{j=1}^{k} N_{j}$ be the $kth$ partial sum of the $N$'s, for $1 \leq k \leq 2n+1$. The key to the algorithm is that the following two events are the same:
%
    \{X_{(i)} \in (h_{i}, g_{i}) \text{ for } i = 1, ..., n\} = \{l_{k} \leq S_{k} \leq u_{k} \textrm{ for } k = 1, 2, ..., 2n\},
%
where
%

@@

%

@@

% Note, need to change all of the l values below to m's to match what the algorithm says above
% Also need to finish writing out the algorithm
% I can do all of this either tonight or tomorrow at some point
% Then I also need to get the bibliography and to the pre-computation for the graphs.

@@

Empirically, it seems that our algorithm has computational complexity $O(n^{2})$. For a dense grid of values of $n$ between $10$ and $50,000$, the number of recursive steps required (line 15 in algorithm 1 above) scales with $n^{2}$. More precisely, it appears that the number of recursive steps required is very close to $8n^{2}$. While each recursive step itself requires calculating a binomial probability which is $O(n)$ due to the calculation of the binomial coefficient multiplied by a quantity with powers as large as $n$, these calculations can be memoized with $O(n^{2})$ cost. Thus, these binomial probability calculations do not change the big $O$ complexity of the algorithm.

@@

As noted above, the algorithm presented is suitable for the case when all of the parameters of $F_{0}$ are known, and all $X_{1}, ..., X_{n}$ are independent. While these assumptions may seem restrictive, the methods above are still useful in a variety of scenarios.

@@

One of the main advantages of the local levels method compared to other global testing approaches, is that it can easily be used to put testing bands onto QQplots by simply graphing each $(h_{i}, g_{i})$ interval. This allows us to examine how a dataset might deviate from some null distribution much better than simply applying a test that yields a binary conclusion. Below, we present a few examples where a QQplot is useful, and where the local levels test seems ideal for assessing deviation from a global null hypothesis.

@@

% It may also make sense for me to make a comparison to the Kolmogorov-Smirnov statistic at some point. I should probably include some instructive graphs or something.
% I want to have a different paragraph to better introduce the concept
% This shouldn't be too difficult
Suppose we have devised a new testing procedure to test a null hypothesis $H_{0}$ with test statistic $T$, where we also specify a particular method to calculate or approximate p-values. In such a situation it is important to perform some simulations under the null hypothesis and check that the resulting p-value distribution is approximately uniform in the simulation experiment.
\newline
\newline
Typically, the verification of Type I error rate is done using the following procedure:
\newline
\newline
(1) Generate $n$ simulated datasets under $H_{0}$, and calculate $T$ for each siulated dataset to obtain $T_{1}, ..., T_{n}$.
\newline
\newline
(2) Select a value of $\alpha$, and for each of $T_{1}, ..., T_{n}$, determine whether the null hypothesis is rejected at level $\alpha$. Let $N_{\alpha}$ be the observed number of the $n$ tests that are rejected at level $\alpha$.
\newline
\newline
(3) Let $\alpha^{*}$ denote the true probability of rejection under the above procedure. Test the null hypothesis $H_{0}: \alpha^{*} = \alpha$ by applying, e.g., a z-test of proportions or an exact binomial test to the data $N_{\alpha}$.
\newline
\newline
While the above procedure provides reliable information about the Type I error calibration for one level of $\alpha$, it provides little information about the global calibration of p-values. Instead, we suggest the following procedure:
\newline
\newline
(1) As above.
\newline
\newline
(2) For each $T_{i}$, calculate the corresponding p-value, $p_{i}$, to obtain $p_{1}, ..., p_{n}$.
\newline
\newline
(3) Make a q-q plot comparing $p_{1}, ..., p_{n}$ to a $U(0, 1)$ distribution, and apply the local levels procedure to create simultaneous testing bands for the null hypothesis that $p_{1}, ..., p_{n} \distras{iid} U(0, 1)$.
\newline
\newline
This allows us to easily visualize the global calibration of the p-values with just one graph and diagnose any issues if they exist. In Step 3, one could use many different testing bands. However, in the calibration of p-values, we typically don't have the expectation that our p-values would be more likely to deviate from uniform in any particular region, and so it makes sense to use the local levels test because it is agnostic to the alternative distribution. Moreover, since it is generally most concerning if small p-values are not calibrated, i.e. those in the lower tail of the uniform distribution, the local levels test is preferable to the standard KS test because it is much more sensitive in the tails (Aldor-Noiman 2013).

\subsubsection{Chi-Square Test for Independence in a $2$ X $2$ Table}
We apply this approach to assess the calibration of p-values from the Pearson Chi-Square test for independence in a $2 \times 2$ table.  A well-known rule of thumb is that the chi-square test is appropriate as long as the expected cell count in each cell under the null hypothesis is at least 5.  We fix the cell probabilities under the null hypothesis and consider two different cases: in scenario 1, the sample size is $s=200$ and the rule of thumb holds, and scenario 2, the sample size is only $s=20$ and the rule of thumb does not hold.  We use the local levels approach to generate simultaneous testing bands to assess the calibration of the p-values from the Pearson Chi-Square test for these two scenarios.\\

More specifically, in each scenario, we randomly generate $n=1000$ $2 \times 2$ tables under the null hypothesis, where each table contains $s$ observations, with $s=200$ in scenario 1 and $s=20$ in scenario 2.  For each table, the $s$ observations are i.i.d.\ with probability $q_{i,j}$ of falling in cell $(i,j)$, for $i=0,1, j=0,1$, where $q_{1,1} = a*b$, $q_{1,0} = a*(1-b)$, $q_{0,1} = (1-a)*b$, and $q_{0,0} = (1-a)*(1-b)$, with a=.15 and b=.4.  For each table, let $X_{i,j}$ denote the observed count in cell $(i,j)$.  If any table has $X_{0,0}+X_{0,1} = 0$ or $= s$, we discard the table and draw a new one, because that would imply that one of the rows of the table is empty, in which case the Pearson Chi-Squared test statistic is not defined.  Similarly, if any table has $X_{0,0}+X_{1,0} = 0$ or $=s$, we discard the table and draw a new one. The code to do this is shown below:
%
<<contingency_tables>>=
set.seed(8675309)
# Generate contingency tables for variables as defined above
generate_contingency_tables <- function(n, s, a, b) {

  cell_probs <- c(a * b, a * (1 - b), b * (1 - a), (1 - b) * (1 - a))
  num_sims <- 0
  sim_mat <- matrix(NA, nrow = n, ncol = 4)

  # Generate n simulations and append the contingency tables to a matrix
  while (num_sims < n) {

    proposed_sim <- t(rmultinom(n = 1, size = s, prob = cell_probs))

    # Check to make sure that the chi-square statistic will be defined
    if (proposed_sim[, 1] + proposed_sim[, 2] != 0 && 
        proposed_sim[, 1] + proposed_sim[, 3] != 0 &&
        proposed_sim[, 3] + proposed_sim[, 4] != 0 &&
        proposed_sim[, 4] + proposed_sim[, 2]) {

      num_sims <- num_sims + 1
      sim_mat[num_sims, ] <- proposed_sim

    }

  }

  return(sim_mat)

}

scenario_1 <- generate_contingency_tables(n = 1000, s = 200, a = .15, b = .4)
scenario_2 <- generate_contingency_tables(n = 1000, s = 20, a = .15, b = .4)
@
%

For each of the tables in the resulting sample, a Pearson chi-square test statistic is calculated as $T = \sum_{i=0}^1\sum_{j=0}^1 \frac{(X_{i,j}-q_{i,j})^2}{q_{i,j}}$, where $X_{i,j}$ is the observed count in cell $(i,j)$.  For each scenario, this results in $n=1000$ test statistics, $T_1, \ldots, T_n$, one for each table.  From these, we obtain $n=1000$ p-values, $p_1, \ldots, p_n$ by applying the $\chi^2_1$ approximation, i.e., $p_i = 1-F(T_i)$ for $i=1,\ldots, n$, where $F$ is taken to be the c.d.f. of the $\chi^2_1$ distribution. The code to do this is shown below:

%
<<p_values>>=
# Compute Pearson Chi-Square Test p-values for Set of Contingency Tables
get_p_values_from_c_tables <- function(c_tables, s) {

  chisq_numerator <- (
    (c_tables[, 1] * c_tables[, 4] - c_tables[, 2] * c_tables[, 3]) / s
  ) ^ 2
  chisq_denominator <- (c_tables[, 1] + c_tables[, 2]) * 
    (c_tables[, 1] + c_tables[, 3]) * 
    (c_tables[, 3] + c_tables[, 4]) * 
    (c_tables[, 2] + c_tables[, 4]) / 
    (s ^ 3)
  chisq_stats_vec <- chisq_numerator / chisq_denominator
  p_vals <- pchisq(q = chisq_stats_vec, df = 1, lower.tail = FALSE)
  return(p_vals)

}

pvals_scenario_1 <- get_p_values_from_c_tables(scenario_1, s = 200)
pvals_scenario_2 <- get_p_values_from_c_tables(scenario_2, s = 20)
@
%

Figure 3 shows the resulting QQplots for scenarios 1 (in blue) and 2 (in red), where the 45$^{\mbox{\footnotesize o}}$ line is shown as well as the testing bands obtained from the equal local levels procedure for testing, at global level .05, the null hypothesis that $p_1, \ldots, p_n$ have the same distribution as $n$ i.i.d.\ draws from Unif(0,1).

<<chisq_qq, fig=TRUE, include=FALSE>>=
library(qqconf)
bounds <- get_bounds_two_sided(alpha = .05, n = 1000)

# Plot 3: Q-Q plot of P-values from Chi-Squared Test
# for Independence in a 2 x 2 Table

plot(
  bounds$x, 
  sort(pvals_scenario_2), 
  type = "l", 
  col = "red", 
  xlab = "Expected Uniform Quantile",
  ylab = "Observed P-value"
)

# Add the 45-degree line to the plot
lines(bounds$x, bounds$x)

# Add the lower bound line for the level .05 
# simultaneous testing bands from equal local levels.
lines(bounds$x, bounds$lower_bound)

# Add the upper bound line for the level .05 
# simultaneous testing bands from equal local levels.
lines(bounds$x, bounds$upper_bound)

# Add the Q-Q plot for scenario 1.
lines(bounds$x, sort(pvals_scenario_1), col = "blue")

legend(
  x = .02, 
  y = .95, 
  legend = c("s=20","s=200"), 
  col = c("red","blue"), 
  lty = 1
)
@

\begin{figure}
\begin{center}
\includegraphics{manuscript-chisq_qq}
\end{center}
\caption{QQplot for Scenario 1 and Scenario 2 With Standard Axes.}
\end{figure}

Figure 4 shows the same plots where the axes are transformed to be on the $-\log_{10}$ scale to focus the plot on small p-values. (Note that in Figure 4, small p-values are to the top and right of the plot, so a curve that is too low is conservative, and too high is anti-conservative.)

<<chisq_qq_log, fig=TRUE, include=FALSE>>=
# Plot 4: Q-Q plot, on the -log10 Scale, of P-values 
# from Chi-Squared Test for Independence in a 2 x 2 Table

plot(
  -log10(bounds$x), 
  -log10(sort(pvals_scenario_2)),
  type = "l",
  col = "red",
  xlab = "-log10(Expected Uniform Quantile)",
  ylab = "-log10(Observed P-value)"
)

# Add the 45-degree line to the plot
lines(-log10(bounds$x), -log10(bounds$x))

# Add the lower bound line for the level .05 
# simultaneous testing bands from equal local levels.
lines(-log10(bounds$x), -log10(bounds$lower_bound))

# Add the upper bound line for the level .05 
# simultaneous testing bands from equal local levels.
lines(-log10(bounds$x), -log10(bounds$upper_bound))

# Add the Q-Q plot for scenario 1, with axes on the -log10 scale.
lines(-log10(bounds$x), -log10(sort(pvals_scenario_1)), col = "blue")

legend(
  x = .2, 
  y = 3, 
  legend = c("s=20", "s=200"), 
  col = c("red","blue"), 
  lty = 1
)
@

\begin{figure}
\begin{center}
\includegraphics{manuscript-chisq_qq_log}
\end{center}
\caption{QQplot for Scenario 1 and Scenario 2 With Axes On Log Scale.}
\end{figure}

From these plots, it can be seen that in scenario 1, when $s=200$ and the smallest expected cell count is 12, there is no significant deviation of the p-values from i.i.d.\ Unif(0,1) under the null hypothesis.  In contrast, in scenario 2, when $s=20$ and the smallest expected cell count is 1.2, the $\chi^2_1$ asymptotic distribution is not an accurate approximation to the sampling distribution of $T$.  As a result, we can see in Figures 3 and 4 that the p-values differ significantly from i.i.d.\ Unif(0,1) under the null hypothesis, with small p-values tending to be overly conservative, while the larger p-values tend to be anti-conservative.

\subsection{Multiple Testing in Genome Wide Association Studies}
In genome wide association studies (GWAS), we are interested in identifying the genetic variants that are responsible for biological traits. Often, we test a large number of single-nucleotide polymorphisms (SNPs), to see if any of them are strongly correlated with some discrete or continuous trait. Often, it is useful to test the following global null hypothesis:
$H_{0}$: All genome-wide SNPs in some functional category have correlation 0 with the trait of interest.
%
\textrm{ vs. }
%
\noindent$H_{A}$: At least one SNP in some functional category has non-zero correlation with the trait of interest.
Typically, when testing this global null, we apply some test to each of $n$ SNPs, yielding $n$ p-values. While the SNPs have a local correlation structure, this has only a weak effect at a genome-wide scale. Thus, the local levels procedure still provides a valuable visualization tool to assess the extent and type of any deviation from the null, even though it is not exact. Particularly in genetics, if the p-values deviate from the null, it can be very useful to view graphically how they deviate. For instance, if there are a few very large p-values but the rest of the distribution looks relatively uniform, then this indicates that the trait of interest is driven by a few mutations. If, however, there are some small deviations throughout the p-value distribution, then this indicates that trait of interest is driven by a number of genes that all play some small part in a complex biological process, or potentially by some confounding variables that are not controlled for.
If a very specific distribution for $H_{A}$ is suspected, then it makes sense to use a test that has high power under this alternative, which may not be our local levels test. However, in the general case when little is known about the alternative distribution, using local levels is desirable because it is agnostic to alternative distributions. Moreover, since small p-values are often of great interest in GWAS because they can indicate the main biological drivers of a trait, the local levels test is far superior to the KS test due to its tail sensitivity.
%Here is where the biological example goes with the code. I want to have the graphs that I make from this and then I also want
\section{Discussion}
The local levels test using an exact recursive calculation allows for the creation of relatively fast and alternative-agnostic simultaneous testing bands for QQplots. Of course, our method is limited by the assumptions that (1) all of the tests are independent and that (2) all of the parameters of the null distribution are known. Currently, we are not aware of methods to handle the dependent case that allow for bands on QQplots. To handle the second issue, Aldor-Noiman has adapted their method such that it will yield exact results when some parameters of the null distribution are unknown. However, if the dataset is very large, Aldor-Noiman's method is prohibitively slow. In this case, we suggest that our method be used because it is much faster, and with a sufficiently large dataset the uncertainty in the parameters will become negligible under standard regularity conditions. Methods to handle parameter uncertainty with an exact calculation have been discussed in the context of the normal distribution \citep{rosenkrantz2000confidence}, but a general method towards this end has not been developed, and would be an important next step.
Our software is available on CRAN. To aid in computation speed, we have pre-computed a number of different local level values given a global level and the size of the dataset. This computes the testing bands very quickly with little error. We have also implemented a one-sided test for the case in which we are only concerned with values that are smaller than would be expected under the null (e.g. we only care about small p-values).

\bibliography{refs}
%\bibliographystyle{ieeetr}

\end{document}
